---
title: "Cocoa Percentage in Chocolate"
author: "Sheli Feigin, Daniel Yohan"
id: "Sheli - 207456302, Daniel - 207106493"
output: html_document
date: '2022-05-29'
editor_options: 
markdown: 
wrap: sentence
---

## Background

Who doesn't like chocolate?

We very much do, and believe (almost) everyone too.
Therefore, we find that the interesting questions should be asked about the preferences people have regarding chocolate. 

In this markdown, we'll analyze the chart from the following site: <http://flavorsofcacao.com/chocolate_database.html>

## Goals

In our research we'll focus on 2 main questions:

1.  Can we assume there's a linear regression between the cocoa percentage (as an independent variable) type of ingredients (as "dummies"), and the chocolate rating (as a dependent variable)?

2.  Are the percentage of cocoa, and the manufactures' location (in which continent they're based) are independent from each other?


### Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(ggplot2)
```

### Importing the Dataset

```{r data import}
cocoa_data <- read_excel("cocoa_data.xlsx") 
```

### Early Filtering and Tidying

The data set is extracted from an Excel file, imported from the linked site above.
We've dropped columns that are irrelevant to our tests: 'REF' and 'Most Memorable Characteristics'.
Likewise, we've omitted rows with missing data, and the given cocoa percentages had to be converted to actual numbers.

```{r early filtering}
drop <- c("REF", "Most Memorable Characteristics")
cocoa_data_clean = cocoa_data[,!(names(cocoa_data) %in% drop)]
cocoa_data_clean <- na.omit(cocoa_data_clean)
cocoa_data_clean[6] <- 100 * (cocoa_data_clean[6])
```

Some header names had to be changed:

```{r headers corrections}
names(cocoa_data_clean)[1] <- 'Company_(Manufacturer)'
names(cocoa_data_clean)[2] <- 'Company_Location'
names(cocoa_data_clean)[3] <- 'Review_Year'
names(cocoa_data_clean)[4] <- 'Bean_Origin'
names(cocoa_data_clean)[5] <- 'Bean_Origin_and_or_Bar_name'
names(cocoa_data_clean)[6] <- 'Cocoa_percent'
```

## First test

In a significance level of 5%:

Our Null hypothesis is that all of the coefficients equal to 0, which means that we can't state there's a linear regression between any of the variables above.

Our Alternative hypothesis is that at least one coefficients isn't 0, which means that we can approve that there's a linear regression between some of the variables.

## Transformation, Modeling and Visualization

### Transformation

We've added 7 columns for each ingredient, indicating whether it's in the chocolate bar or not. Then we extracted the types of ingredients of each bar and filled the new columns accordingly. 

```{r test1 transformation}
cocoa_data_clean[7] <- gsub("^.*?- ", "", cocoa_data_clean$Ingredients)
cocoa_data_clean[c("Beans", "Sugar", "Sweetener", "Cocoa_Butter", "Vanilla", "Lecithin", "Salt")] <- 0
cocoa_data_clean$Ingredients <- strsplit(cocoa_data_clean$Ingredients, split = ",")
for (i in 1:length(cocoa_data_clean$Ingredients)) {
  curr_ingredients <- cocoa_data_clean$Ingredients[[i]]
  if ("B" %in% curr_ingredients) {
    cocoa_data_clean$Beans[i] <- 1
  }
   if ("S" %in% curr_ingredients) {
    cocoa_data_clean$Sugar[i] <- 1
   }
   if ("S*" %in% curr_ingredients) {
    cocoa_data_clean$Sweetener[i] <- 1
   }
   if ("C" %in% curr_ingredients) {
    cocoa_data_clean$Cocoa_Butter[i] <- 1
   }
   if ("V" %in% curr_ingredients) {
    cocoa_data_clean$Vanilla[i] <- 1
   }
   if ("L" %in% curr_ingredients) {
    cocoa_data_clean$Lecithin[i] <- 1
   }
   if ("Sa" %in% curr_ingredients) {
    cocoa_data_clean$Salt[i] <- 1
   }
}

```

### Visualization

Let's get a first impression regarding the cocoa percentage and ratings in out data set:

```{r test1 visualzation 1}

cocoa_data_clean %>%
ggplot(aes(x = Cocoa_percent, y = Rating)) +
  geom_jitter(alpha = .75) + 
  coord_cartesian(ylim = c(0,5)) +
  labs(x = 'Cocoa Percentage', y = 'Rating') + 
  theme_minimal() + 
  geom_smooth(method = 'lm', se = FALSE, col = 'chocolate')

```

We can't notice a strong relationship between the cocoa percentage and the chocolates' rating.

```{r test1 visualzation 2}
hist(cocoa_data_clean$Cocoa_percent,
     col = "chocolate",
     main = "Cocoa Percent Histogram",
     xlab = "Cocoa Percent")

data_mean <- mean(cocoa_data_clean$Cocoa_percent)
data_sd <- sd(cocoa_data_clean$Cocoa_percent)
x <- seq(40, 100, length=nrow(cocoa_data_clean))
y <- dnorm(x, mean = data_mean, sd = data_sd) * 5 * nrow(cocoa_data_clean)
lines(x, y, lwd = 2)
```

Here we can roughly see that the frequency of the cocoa percentage distribute normally. Moreover, most of the documented chocolate bars consist between 65% - 75% cocoa.

Note: in order to convert the normal distribution plot from density to frequency, we had to multiply its y values by the histogram bars width (5) and the number of observations (nrow). 

```{r test1 visualzation 3}
hist(cocoa_data_clean$Rating,
     col = "chocolate",
     main = "Chocolate Rating Histogram",
     xlab = "Rating",
     breaks = seq(1.0, 4.0, 0.25),
     xaxp = c(1.0, 4.0, 12)
     )
```

From the second Histogram we understand that the vast majority of the chocolate bars have a rating greater than 2.25.

### Modeling

Here we've modeled the linear regression test:

```{r test1 model table}
rating <- cocoa_data_clean$Rating
cocoa_percent <- cocoa_data_clean$Cocoa_percent
sugar <- cocoa_data_clean$Sugar
sweetener <- cocoa_data_clean$Sweetener
cocoa_Butter <- cocoa_data_clean$Cocoa_Butter
vanilla <- cocoa_data_clean$Vanilla
lecithin <- cocoa_data_clean$Lecithin
salt <- cocoa_data_clean$Salt

model <- lm( formula = rating ~ cocoa_percent + sugar + sweetener + cocoa_Butter + vanilla + lecithin + salt, data = cocoa_data_clean)
model %>% summary()

```
We can observe that as the cocoa percentage gets higher the chocolates' rating slightly decreases.
Moreover, the presence of sweeteners, vanilla, lecithin and salt also decrease the rating - whereas cocoa butter and lecithin increase the rating.  

In the following chunk we'll examine if homoscedasticity exists in our model:

```{r test1 homoscedasticity}
res <- model$residuals
fit <- model$fitted.values
res_table= tibble(res,fit)

ggplot(res_table,aes(x=res)) + geom_density(color="chocolate") + xlab("Residuals") + ggtitle("Residuals Distribution")

scatter.smooth(x=fit, y=res, main="Homoscedasticity", xlab = "Fitted Values", ylab = "Residuals")

ggplot(res_table, aes(sample=res)) + geom_qq() + geom_qq_line(color="chocolate") + xlab("Observed Value") + ylab("Expected Normal") 


```

In the first plot we see that the residuals don't distribute normally.
We notice from the second plot that the variance is not quite consistent along the fitted values.
Moreover, it's seen from the third that the residuals can't be found on the line - also indicating they don't distribute normally.
Consequently, we infer there's no homoscedasticity in our model.
In any case we'll conclude the test as we could assume homoscedasticity.

### Conclusion

Our PV is smaller then 5%, therefore we deny the null hypothesis, and accept the alternative.
We approve there's a linear connections between some of the variables.

## Second Test

In a significance level of 5%:

Our Null hypothesis is that the percentage of cocoa and the continents of the chocolate manufacturers are independent variables.

Our Alternative hypothesis is that those variables are dependent.

## Transformation, Modeling and Visualization

### Transformation

The transformation for this test was more complex.

Firstly, we've created vectors of continents for each chocolate companies' location included in the data set.

```{r continent vectors}
in_africa = c("Sao Tome",	"Ghana",	"Madagascar",	"South Africa") 
in_asia = c("Japan",	"India",	"Israel",	"Malaysia",	"Philippines",	"Russia",	"Singapore",	"South Korea",	"Taiwan",	"Thailand",	"U.A.E.",	"Vietnam")
in_australia = c("Australia",	"Fiji",	"New Zealand",	"Vanuatu")
in_europe = c("U.K.",	"France",	"Germany",	"Italy",	"Amsterdam",	"Austria",	"Belgium",	"Spain",	"Czech Republic",	"Denmark",	"Finland",	"Hungary",	"Iceland",	"Ireland",	"Lithuania",	"Netherlands",	"Norway",	"Poland",	"Portugal",	"Scotland",	"Sweden",	"Switzerland",	"Wales", "8")
in_north_america = c("Canada",	"U.S.A.",	"Costa Rica",	"Dominican Republic",	"El Salvador",	"Grenada",	"Guatemala",	"Honduras",	"Martinique",	"Mexico",	"Nicaragua",	"Puerto Rico",	"Sao Tome 0%_to_60%& Principe",	"St. Lucia",	"St.Vincent-Grenadines")
in_south_america = c("Ecuador",	"Argentina",	"Bolivia",	"Brazil",	"Chile",	"Colombia",	"Peru",	"Suriname",	"Venezuela")
```

Then we've made a new table with observations as an input for the Chi-square test.

```{r observed table}
observed_table <- matrix(0, ncol = 4, nrow = 6)
continent_names <- c("Africa", "Asia", "Australia", "Europe", "North America", "South America")
rownames(observed_table) <- continent_names
colnames(observed_table) <- c("40%-66%", "67%-74%", "75%-80%", "81%-100%")

```

In the following chunk we've added the Continent column.
We also created a for-loop that filled each row in the main data set with its' appropriate continent, and simultaneously filled our observed_table for the square chi test.

```{r filling loop}
cocoa_data_clean$Continent <- NA
for (i in 1:length(cocoa_data_clean$Company_Location)) {
  if (cocoa_data_clean$Company_Location[i] %in% in_africa) {
    cocoa_data_clean$Continent[i] <- "Africa"
    fill_row = 1
  }
  else if (cocoa_data_clean$Company_Location[i] %in% in_asia) {
    cocoa_data_clean$Continent[i] <- "Asia"
    fill_row = 2
  }
  else if (cocoa_data_clean$Company_Location[i] %in% in_australia) {
    cocoa_data_clean$Continent[i] <- "Australia"
    fill_row = 3
  }
  else if (cocoa_data_clean$Company_Location[i] %in% in_europe) {
    cocoa_data_clean$Continent[i] <- "Europe"
    fill_row = 4
  }
  else if (cocoa_data_clean$Company_Location[i] %in% in_north_america) {
    cocoa_data_clean$Continent[i] <- "North America"
    fill_row = 5
  }
  else {
    cocoa_data_clean$Continent[i] <- "South America"
    fill_row = 6
  }
  if (cocoa_data_clean$Cocoa_percent[i] %in% seq(40,68)) {
  observed_table[fill_row, 1] = observed_table[fill_row, 1] + 1
  }
  else if (cocoa_data_clean$Cocoa_percent[i] %in% seq(69,74)) {
  observed_table[fill_row, 2] = observed_table[fill_row, 2] + 1
  }
  else if (cocoa_data_clean$Cocoa_percent[i] %in% seq(75,80)) {
  observed_table[fill_row, 3] = observed_table[fill_row, 3] + 1
  }
  else {
  observed_table[fill_row, 4] = observed_table[fill_row, 4] + 1
  }
}
```

### Visualization

We wanted to take a look at the connection of cocoa percentage and the manufacturers' continent, so we've created box-plots graph to see how the cocoa percentages are scattered in each continent.

```{r test2 visualzation}
boxplot(cocoa_data_clean$Cocoa_percent ~ cocoa_data_clean$Continent,
data=cocoa_data_clean,
main="Different boxplot for each continent",
xlab="Continents",
ylab="Cocoa precent",
col="chocolate",
border="brown",
pch = 19,
cex.axis = 0.8
)
rowSums(observed_table)
```

We can see that the continents have a similar median values of cocoa percentage - around 70%.
The exception is Africa, with a median close to 75%. Worth mentioning that the continent doesn't have many observations - so its median may be more biased compared to others.    

The two middle quarters of each continent are between 70% and 75%.

A noticeable difference between the boxes is within the extreme values.
The continents with relatively more occurrences in the data set (Europe and North America) have a greater number of extremes, which makes sense. 

### Modeling

Let's examine the expected values for our independence test.

```{r check expected}
chi_test <- chisq.test(observed_table)
chi_test$expected
```

There are cells for Africa and Australia with expected data smaller then 5.
Therefore, we've merged the data of the two continent into one.

```{r merging continents}
for (i in 1:4) {
  observed_table[1, i] <- observed_table[1, i] + observed_table[3, i]
}
fixed_observed_table = observed_table[-3,]
rownames(fixed_observed_table)[1] <- "Africa & Australia"
```

Now each cell has a value greater then 5 and we can properly conduct our second test.

```{r check fixed expected}
chi_test <- chisq.test(fixed_observed_table) 
chi_test$expected
chi_test
```

### Conclusion

Our PV is smaller then 5%.
In that significance level, we deny the null hypothesis and accept the alternative.
Therefore, we can state that the percentage of cocoa, and the continents of the chocolate manufacturers are indeed dependent.
